---
title: "Study case of three crop scenarios - landscape metrics"
author: "team_cli07_landscapeoptimizer"
date: "12 July 2020"
output:
  html_document:
    toc: true
    toc_float: true
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=TRUE)
```

```{r}
library(raster)
library(sf)
library(landscapemetrics)

library(dplyr)
library(ggplot2)
library(mapview)
```

# Scenarios

The user would be able to generate several scenarios by assigning different crops to their agricultural patches.

```{r, read-data}
# Read rasters with crop data - to be used directly for landscapemetrics::sample_lsm()
rst <- list.files(path = "data", pattern = "\\.tif", full.names = TRUE) %>% 
  lapply(FUN = raster)

# Read polygons with crop data
polys <- read_sf("data/AOIParcelsWithInfo.shp")

# Read biogas plant location
load(file = "data/biogas.plant.sf.rda")
pts <- as(biogas.plant.sf, Class = "Spatial")
```

Below one can see three different crop scenarios. Scenario 2 is an extreme example of when an user would decide to plant one single type of species on her/his agricultural patches (monocultures).
```{r, plot-crops}
polys <- polys %>% select(land_cover = LndCvrT,
                          matches("Crops"))
mapview(polys, zcol = "Crops1")
mapview(polys, zcol = "Crops2")
mapview(polys, zcol = "Crops3")
```

# Landscape diversity metrics

The user will get statistics about some relevant landscape metrics like edge density which is correlated to species diversity. 

The higher the edge density and patch density, the higher the species diversity within the area of study.

The Shannon diversity index indicates the diversity of the landscape. In this case, scenario 3 has the highest diversity index. Scenario 2 (maze only) surprisingly performs better than scenario 1.
```{r, get-metrics}
# Compute landscape metrics of interest
landscape_metrics_available <- list_lsm(level = "landscape")
metrics_of_interest <- c("lsm_l_ed", "lsm_l_pd", "lsm_l_shdi")

metrics <- sample_lsm(landscape = rst,
                      y = pts, 
                      shape = "square",
                      size = 2500,
                      what = metrics_of_interest) %>% 
  left_join(y = landscape_metrics_available,
            by = "metric")
```

```{r, plot-metrics}
ggplot(data = metrics,
       aes(x = factor(layer),
           y = value)) +
  geom_col(fill = "gray70") +
  facet_wrap(~ name, scales = "free") +
  labs(x = "Scenario") +
  theme_bw()
```

