---
title: "Demo landscape metrics on a raster dataset prepared by Niko"
author: "Valentin Stefan"
date: "First created: 11 July 2020; Last update: `r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=TRUE)
```

```{r}
# for basic example
library(raster)
library(landscapemetrics)
# for some more fancy example
library(dplyr)
library(purrr)
library(ggplot2)
library(plotly)

library(knitr)
library(kableExtra)
```


# Landscape metrics for given location and buffer size

The raster must have the land use classes encoded as integers and must be projected in a metric coordinate system (meters, not degrees).

General checkup if the raster fulfill the conditions to be used (e.g.: is it projected, does it have a CRS, etc.):

```{r}
rst <- raster("LandCoverParcelsFasterized.tif")
# rst <- raster("./sandbox/LandCoverParcelsFasterized.tif")
crs(rst) <- CRS('+init=EPSG:32632')
```

Make a fake point (a hypothetical location for a biomass plant). A circular buffer will be used to compute the landscape metrics within that buffer. Can also be multiple points and also can be a spatial object not only a matrix of projected coordinates:
```{r}
pts <- matrix(data = c(720000, 5630000), ncol = 2)

plot(rst); points(pts, cex = 2, pch = 16)
```

Get landscape metrics as data frame:
```{r}
landscape_metrics_available <- list_lsm(level = "landscape")

metrics_of_interest <- c("lsm_l_ed", "lsm_l_te", "lsm_l_ta",
                         "lsm_l_area_mn", "lsm_l_area_sd", "lsm_l_area_cv", 
                         "lsm_l_pd", "lsm_l_shdi")

metrics <- sample_lsm(landscape = rst,
                      y = pts, 
                      shape = "square", # buffer shape
                      size = 5000,      # meters - half of the side-length of the buffer
                      what = metrics_of_interest) 

metrics <- left_join(x = metrics,
                     y = landscape_metrics_available,
                     by = "metric")
knitr::kable(metrics) %>% kable_styling()
```

# Metrics for several buffer sizes

Try several buffer sizes and visualize interactively how the metrics vary with that.

```{r}
sizes <- seq(from = 10^3, to = 10^4, by = 10^3)

metrics_radii <- sizes %>% 
  set_names() %>% 
  map_dfr(~ sample_lsm(landscape = rst,
                       y = pts, 
                       shape = "square",
                       size = .,
                       what = metrics_of_interest), 
          .id = "sizes")

metrics_radii_4_ggplot <- left_join(x = metrics_radii,
                                    y = landscape_metrics_available,
                                    by = "metric") %>% 
  mutate(sizes = as.integer(sizes),
         facet = paste(name, metric, sep = " - "))
```

Make interactive plot:
```{r}
metrics_plot_radii <- metrics_radii_4_ggplot %>% 
  ggplot(aes(x = sizes,
             y = value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ facet, scales = "free") +
  labs(x = "Side-length of square buffer (m)") +
  theme_bw()

ggplotly(metrics_plot_radii)
```